Moisture effects on leaf litter are species-specific and result in non-additive flammability in mixed conifer forest

### 2019-07-24 ms/Magalhaes_Schwilk_litter_flam_moisture.tex
edit abstract

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{ Moisture content is a strong determinant of forest fuel flammability
  and varies on small spatial scales and over short time periods. Emperical
  modeling of equilibrium dead fine fuel moisture is well established, but less
  is known about the mechanisms by which species-specific litter traits
  influence litter moisture dynamics, particularly the processes of moisture
  absorption and desorption. Characterizing how species’ leaf litter retain and
  release moisture through time is critical to the assessment of fuel’s
  availability to burn. Furthermore, Natural forests are often comprised of
  mixed stands, leading to mixtures of leaf litter on the surface fuelling the
  fires. Multiple studies have established the existence of non-additive
  effects in the flammability of dry litter, but we lack a mechanistic
  understanding of moisture synergies in mixed litter and of moisture effects
  on the flammability of mixed litter. The litter from eight species of a
  mixed-conifer forest was saturated and allowed to dry to determine moisture
  absorption capacity and desorption rates. Burn trials were performed in moist
  litter beds of single species and of mixtures, to establish flammability
  response to the dry-down process. Moisture dynamics vary across species with
  higher specific leaf area being associated with higher maximum absorption and
  lower litter bulk density being associated with faster drying rates. Species
  that produce more aerated litter beds (such as \emph{Q. kelloggii} and
  \emph{Pinus} species) have faster rates of litter drying. We examined two
  axes of flammability (spread rate and fuel consumption) and found that these
  differing moisture dynamics result in time since wetting having strongly
  differing effects on flammability across the major litter types typified by
  different tree genera. Finally, we found that litter mixtures exhibited
  consistent non additivity in flame spread rate in fuel consumption. This non
  additivity became more positive with fuel drying. Litter mixtures burn with
  fire behavior more similar to that of the most flammable constituent species
  and this effect increases as fuels dry.}

@@

[CONCLUSIONS ?]

>>>>

%--------------------
% Class Options:
%--------------------

%----------
% journal
%----------
%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{ Moisture content is a strong determinant of forest fuel flammability
  and varies on small spatial scales and over short time periods. Emperical
  modeling of equilibrium dead fine fuel moisture is well established, but less
  is known about the mechanisms by which species-specific litter traits
  influence litter moisture dynamics, particularly the processes of moisture
  absorption and desorption. Characterizing how species’ leaf litter retain and
  release moisture through time is critical to the assessment of fuel’s
  availability to burn. Furthermore, Natural forests are often comprised of
  mixed stands, leading to mixtures of leaf litter on the surface fuelling the
  fires. Multiple studies have established the existence of non-additive
  effects in the flammability of dry litter, but we lack an understanding of
  moisture synergies in mixed litter and of moisture effects on the
  flammability of mixed litter. The litter from eight species of a
  mixed-conifer forest was saturated and allowed to dry to determine moisture
  absorption capacity and desorption rates. Burn trials were performed in moist
  litter beds of single species and of mixtures to establish flammability
  response to the dry-down process. Moisture dynamics vary across species with
  higher specific leaf area being associated with higher maximum absorption and
  lower litter bulk density being associated with faster drying rates. Species
  that produce more aerated litter beds (such as \emph{Q. kelloggii} and
  \emph{Pinus} species) have faster rates of litter drying. We examined two
  axes of flammability variation (spread rate and fuel consumption) and found that these
  differing moisture dynamics result in time since wetting having strongly
  differing effects on flammability across the major litter types typified by
  different tree genera. Finally, we found that litter mixtures exhibited
  consistent non additivity in flame spread rate in fuel consumption. This non
  additivity became more positive with fuel drying. Litter mixtures burn with
  fire behavior more similar to that of the most flammable constituent species
  and this effect increases as fuels dry. }

@@
